---
title: "**Probability and Statistics with R**"
author: "**Mohit Kumar Jain**"
date: "Submission Nov 16-2022 (Wednesday)"
output: pdf_document
---

## Problem 2 : Simulation Study to Understand Sampling Distribution

**Part A**
Suppose 
$X_1,X_2,\cdots,X_n\stackrel{iid}{\sim} Gamma(\alpha,\sigma)$ with pdf as
$$
f(x | \alpha,\sigma)=\frac{1}{\sigma^{\alpha}\Gamma(\alpha)}e^{- x/\sigma}x^{\alpha-1},~~~~0<x<\infty,
$$
The mean and variance are $E(X)=\alpha\sigma$ and $Var(X)=\alpha\sigma^2$. Note that `shape = ` $\alpha$ and `scale = ` $\sigma$.

1. Write a `function` in `R` which will compute the MLE of $\theta=\log(\alpha)$ using `optim` function in `R`. You can name it `MyMLE`
2. Choose `n=20`, and `alpha=1.5` and `sigma=2.2`
     (i) Simulate $\{X_1,X_2,\cdots,X_n\}$ from `rgamma(n=20,shape=1.5,scale=2.2)`
     (ii) Apply the `MyMLE` to estimate $\theta$ and append the value in a vector
     (iii) Repeat the step (i) and (ii) 1000 times
     (iv) Draw histogram of the estimated MLEs of $\theta$.
     (v) Draw a vertical line using `abline` function at the true value of $\theta$.
     (vi) Use `quantile` function on estimated $\theta$'s to find the 2.5 and 97.5-percentile points. 
3.  Choose `n=40`, and `alpha=1.5` and repeat the (2).
4.  Choose `n=100`, and `alpha=1.5` and repeat the (2).
5. Check if the gap between 2.5 and 97.5-percentile points are shrinking as sample size `n` is increasing?

*Hint*: Perhaps you should think of writing a single `function` where you will provide the values of `n`, `sim_size`, `alpha` and `sigma`; and it will return the desired output. 

\newpage

```{r}
library(ggplot2)
#dataframe to store the quantiles for different sample sizes
quantiles=data.frame('X1'=integer(),'X2'=double(),'X3'=double())
count=1 #pointer to the first empty row in 'quantiles' dataframe
colnames(quantiles)=c('Sample size','2.5 percentile', '97.5 percentile')
```

```{r}
f=function(n,sim_size,alpha,sigma){
  mle_vector=c()
  #generating 'sim_size' many random samples and their MLEs
  for (i in 1:sim_size){
    set.seed(2*i) #for reproducibility of results
    data=rgamma(n,alpha,scale=sigma)
    t1=sum(log(data))
    t2=sum(data)
    n=length(data)
    #MyMLE returns the negative log likelihood function which is then minimised,
    #so as to maximise likelihoood, using 'optim'
    #alpha1 and sigma1 are the initial guess for MLEs of alpha and sigma respectively
    MyMLE=function(x){
      alpha_hat=x[1]
      sigma_hat=x[2]
      #return(-sum(log(dgamma(data,alpha_hat,sigma_hat))))
      return(-((alpha_hat-1)*t1-1/sigma_hat*t2-n*log(gamma(alpha_hat))-n*alpha_hat*log(sigma_hat)))
    } 
    alpha1=mean(data)^2/var(data)
    sigma1=var(data)/mean(data)
    x=optim(c(alpha1,sigma1),MyMLE,method='L-BFGS-B',lower=c(0.01,0.01),upper=c(1000,1000))
    mle_vector=append(mle_vector,log(x$par[1])) #only taking the MLE of log(alpha) from the optimisation result
  }
  df=data.frame(mle_vector)
  colnames(df)=c('theta')
  #plotting the histogram of MLEs over different random samples
  p1=ggplot(df,aes(x=theta))+
       geom_histogram(bins=30,colour='grey',fill='red')+
       labs(x='MLEs of theta',y='No. of occurences',title   
            =paste('Sample size=',n))+
      theme(axis.text.x = element_text(angle = 60, hjust = 1))+
      geom_vline(aes(xintercept=log(alpha)))#at true value of theta
  print(p1)
  #populating the quantiles dataframe
  quantiles[count,]<<-c(n,unname(quantile(mle_vector,probs=c(0.025,0.975))))
  count<<-count+1
}
```

```{r}
#function call for different sample sizes
f(20,1000,1.5,2.2)
f(40,1000,1.5,2.2)
f(100,1000,1.5,2.2)
```
```{r}
#gauging the behaviour of percentile points with change in sample size
quantiles['Inter quantile range']=quantiles['97.5 percentile']-quantiles['2.5 percentile']
quantiles
```

## We clearly see that the gap between percentile points shrinks as sample size increases
